---
title: "QC Report: ACT1 & and ACT3 NHW Genotyped SNPs"
author:
- name: Jai Broome
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    keep_md: yes
---

```{r library, message = FALSE}
library(magrittr)
library(dplyr)
library(ggplot2)
library(purrr)
library(knitr)
input <- snakemake@input
params <- snakemake@params
cohorts <- params$cohorts
```

```{r data}
sampmet <- readRDS(input$sampmet)
samp_keep <- readRDS(input$samp_keep)
varmet <- readRDS(input$varmet)
var_keep <- readRDS(input$var_keep)
```

## Sample QC

We apply filters based on four criteria:

* Missingness
* Excess heterozygosity
* Geno/pheno sex mismatch, calculated with `plink --check-sex`
* Phenotype availability

Phenotype availability practically has the largest effect, but we've known that
only a small subset of samples have BPS scores measured. Check the config for
exclusion thresholds for missingness and heterozygosity (we filter out samples
where the heterozygosity is a certain number of SDs away from mean, not an
absolute threshold).

```{r samp_qc_summary}
summarize_samp_counts <- function(x) {
    c(
        total = nrow(x),
        ms_rm = sum(x$ms_rm, na.rm = TRUE),
        htz_rm = sum(x$htz_rm, na.rm = TRUE),
        sexcheck_rm = sum(x$sexcheck_rm, na.rm = TRUE),
        total_excluded = sum(!x$pass, na.rm = TRUE)
    )
}

print_samp_summary <- function(x) {
    summ_out <- data.frame(merged = summarize_samp_counts(x))
    for (i in cohorts) {
        summ_out[[i]] <- summarize_samp_counts(filter(x, cohort == i))
    }
    kable(summ_out) %>% print()
}
```

Sample QC summary: all samples

```{r samp_qc_summary_all, results="asis"}
print_samp_summary(sampmet)
```

Sample QC summary: NP samples only

```{r samp_qc_summary_np, results="asis"}
filter(sampmet, np_measured) %>%
    print_samp_summary()
```

## Variant QC

Summary of variants kept/excluded by missingness or HWE

**ACT1**

```{r a1_variant_summary}
select(varmet$ACT1, hwe_rm, ms_rm) %>% table(useNA = "ifany")
```

**ACT3 NHW**
```{r a3_nhw_variant_summary}
select(varmet$ACT3_NHW, hwe_rm, ms_rm) %>% table(useNA = "ifany")
```